An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2022-06-03 poly-embed-journal/poly-embed-journal.tex
done with sec 4 proof sketch rewrite

To set the stage for Construction \ref{const:eps-thick-link}, we sketch the two main steps in proving Theorem \ref{thm:thickened-separated}: (i) showing that one can produce a link $\psi$ such that $(L,\psi)$ indirectly elicits $\ell$; (ii) ``thickening'' $\psi$ such that it is separated.

For (i), begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$ for two distinct distributions $p,p'\in\simplex$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
Indirect elicitation implies that they do;
we use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that $L$ must indirectly elicit $\ell$, giving a value of $\psi(u)$ for non-embedding points.
In particular, for each such $u$ there exists $r \in \R$ such that any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
Thus if $u \in \Gamma(p) \cap \Gamma(p')$, then there exists $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

\begin{lemma}\label{lem:embed-implies-indirect-elic}
  Let polyhedral loss $L$ embed some discrete loss $\ell$.
  Then $L$ indirectly elicits $\prop\ell$.
\end{lemma}

For (ii), we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
For any $p \in \simplex$, consider the optimal surrogate report set, i.e.\ set of the form $U = \Gamma(p) = \argmin_u \inprod{p}{L(u)}$.
By indirect elicitation, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $\gamma(p) = \argmin_r \inprod{p}{\ell(r)}$.
For $\epsilon > 0$, this condition directly implies separation.

@@

To do so, given $u \in U_\epsilon \cap U'_\epsilon \cap \dots$, we define a \emph{link envelope} $\Psi(u)$ which encodes the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U_\epsilon,U'_\epsilon$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: at least one permitted value for $\psi(u)$ remains.
This claim follows from a geometric result (Lemma~\ref{lem:thick-empty}) that, for all small enough $\epsilon$, a subset of thickenings $U_{\epsilon}$ intersect if and only if the $U$ sets themselves intersect.
When they do intersect, indirect elicitation implies that there exists a permitted choice of link for the intersection of the thickenings.
It is also important that, by Lemma~\ref{lem:polyhedral-range-gamma}, polyhedral surrogates only have finitely many sets of the form $U = \Gamma(p)$.
Together, these observations yield a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

Given the above proof sketch, the following construction is relatively straightforward.
We initialize the link using the embedding points and optimal report sets, then adjust $\Psi$ to narrow down to only legal choices; we then pick from $\psi(u)$ from $\Psi(u)$ arbitrarily.
Theorem \ref{thm:thickened-separated} implies that, for all small enough $\epsilon$, the resulting link $\psi$ is well-defined at all points, and $\epsilon$-separated.
\begin{construction}[$\epsilon$-thickened link] \label{const:eps-thick-link}
  Given a polyhedral $L$ that embeds some $\ell$ via the embedding $\varphi: \Sc \to \reals^d$ for a representative set $\Sc$; and given an $\epsilon > 0$ and a norm $\|\cdot\|$, the \emph{$\epsilon$-thickened link} $\psi$ is constructed as follows.
  Define $\U = \{\Gamma(p) : p \in \simplex\}$.
  % For each $U \in \U$, let $R_U = \{r \in \R : \varphi(r) \in U\}$, the reports whose embedding points are in $U$. \jessiet{Move this sentence below link envelope definition?}
  First, initialize the \emph{link envelope} $\Psi: \reals^d \to 2^{\Sc}$ by setting $\Psi(u) = \Sc$ for all $u$.
  Then for each $U \in \U$, for all points $u$ such that $\inf_{u^* \in U} \|u^*-u\| < \epsilon$, update $\Psi(u) = \Psi(u) \cap R_U$, where $R_U = \{r \in \Sc : \varphi(r) \in U\}$ is the set of target reports that embed into $U$.
  Finally, define $\psi(u) \in \Psi(u)$, breaking ties arbitrarily.
  If $\Psi(u)$ became empty, then leave $\psi(u)$ undefined.
\end{construction}

\subsubsection{A converse to Theorem \ref{thm:thickened-separated}}
Perhaps surprisingly, one can also show that \emph{every} calibrated link from a polyhedral surrogate to a discrete loss is produced by Construction \ref{const:eps-thick-link}.
This result follows from the fact, stated now, that Construction~\ref{const:eps-thick-link} with $\|\cdot\|_\infty$ is exactly enforcing $\epsilon$-separation.
From Theorem~\ref{thm:calibrated-separated}, every calibrated link $\psi$ is therefore output by the construction for some sufficiently small $\epsilon$, in the sense that $\psi$ is one of the valid choices within the link envelope.

\subsubsection{Remarks}
Applying Construction \ref{const:eps-thick-link} enables one to verify the consistency of a proposed link $\psi^*$.
For a given $\epsilon$ and norm $\|\cdot\|$, suppose one follows the routine of Construction~\ref{const:eps-thick-link} until the last step in which values for the link $\psi$ are selected.
Instead, we can simply test whether the proposed link values are contained in the valid choices, i.e., if $\psi^*(u) \in \Psi(u)$ for all $u\in\reals^d$.
If so, then the proposed link $\psi^*$ is calibrated.
\botodo{In general, how do we deal with the $\epsilon$? How do we know what $\epsilon$ to test? I guess we can gradually shrink it and see...}
\raft{A guess for $\epsilon$ was pretty clear from the top-$k$ paper, so yeah I think ``shrink and see'' makes sense.  Also probably okay to not even mention $\epsilon$ here given the last line of the paragraph.}
See \S~\ref{sec:top-k} for an illustration of this test.
On the other hand, if $\psi^*$ cannot be produced from Construction \ref{const:eps-thick-link}, then it cannot be a calibrated link.
This impossibility can be shown, for example, if there exists a point $u$ where $\Psi(u)$ is empty for all $\epsilon > 0$.

@@

>>>>

To set the stage for Construction \ref{const:eps-thick-link}, we sketch the two main steps in proving Theorem \ref{thm:thickened-separated}: (i) showing that one can define a link $\psi$ on all possible optimal points of $L$; (ii) ``thickening'' $\psi$ so that it is separated.

For (i), given the embedding $\varphi: \Sc\to\reals^d$, begin by linking each embedding point back to its original report, so that $\psi(\varphi(r)) = r$.
Now we wish to determine $\psi(u)$ for non-embedding points $u\in\reals^d$ which optimize $L$ for some distribution.
Let $\Gamma = \prop L$ and define $\U = \{\Gamma(p) \mid p\in\simplex\}$ to be the ``range'' of $\Gamma$, i.e., all possible optimal sets.
For each $U\in\U$, we can define $R_U = \{ r\in\Sc \mid \varphi(r) \in U\}$ to be the set of target reports which, by definition of embedding, are $\ell$-optimal when $U$ is $L$-optimal.
We would like to restrict $\psi(U) \subseteq R_U$, so that optimal surrogate reports are mapped to optimal target reports.
The challenge is that we could have a point $u\in U\cap U'$ for two optimal sets $U,U'\in\U$, and \emph{a priori}, it could be that $R_U \cap R_{U'} = \emptyset$.
The first step of the proof is to show that this scenario cannot arise.
\begin{lemma}\label{lem:intersection-property-embedding}
  % Let polyhedral loss $L$ embed some discrete loss $\ell$.
  Let $\U'\subseteq\U$.  Then $\cap_{U\in\U'} U \neq \emptyset$ implies $\cap_{U\in\U'} R_u \neq \emptyset$.
\end{lemma}
\noindent
Thus, for any $u\in\cup\U$, we have a nonempty set of valid choices for $\psi(u)$.
(The statement of Lemma~\ref{lem:intersection-property-embedding} may appear similar to indirect elicitation; in fact the two conditions are equivalent, as we discuss in \S~\ref{sec:poly-ie-consistency}.)

For (ii), we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
Let $U\in\U$.
By the above, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $R_U$.
For $\epsilon > 0$, this condition directly implies separation.

@@

To do so, given $u \in U_\epsilon \cap U'_\epsilon \cap \dots$, we define a \emph{link envelope} $\Psi(u)$ which encodes the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U_\epsilon,U'_\epsilon$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: at least one permitted value for $\psi(u)$ remains.
This claim follows from a geometric result (Lemma~\ref{lem:thick-empty}) that, for all small enough $\epsilon$, a subset of thickenings $U_{\epsilon}$ intersect if and only if the $U$ sets themselves intersect.
When they do intersect, Lemma~\ref{lem:intersection-property-embedding} implies that there exists a permitted choice of link for the intersection of the thickenings.
It is crucial that, by Lemma~\ref{lem:polyhedral-range-gamma}, polyhedral surrogates only have finitely many sets of the form $U = \Gamma(p)$.
Together, these observations yield a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

Given the above proof sketch, the following construction is relatively straightforward.
We initialize the link using the embedding points and optimal report sets, then adjust $\Psi$ to narrow down to only legal choices; we then pick from $\psi(u)$ from $\Psi(u)$ arbitrarily.
Theorem \ref{thm:thickened-separated} implies that, for all small enough $\epsilon$, the resulting link $\psi$ is well-defined at all points, and $\epsilon$-separated.
\begin{construction}[$\epsilon$-thickened link] \label{const:eps-thick-link}
  Given a polyhedral $L$ that embeds some $\ell$ via the embedding $\varphi: \Sc \to \reals^d$ for a representative set $\Sc$; and given an $\epsilon > 0$ and a norm $\|\cdot\|$, the \emph{$\epsilon$-thickened link} $\psi$ is constructed as follows.
  Define $\U = \{\Gamma(p) : p \in \simplex\}$.
  % For each $U \in \U$, let $R_U = \{r \in \R : \varphi(r) \in U\}$, the reports whose embedding points are in $U$. \jessiet{Move this sentence below link envelope definition?}
  First, initialize the \emph{link envelope} $\Psi: \reals^d \to 2^{\Sc}$ by setting $\Psi(u) = \Sc$ for all $u$.
  Then for each $U \in \U$, for all points $u$ such that $\inf_{u^* \in U} \|u^*-u\| < \epsilon$, update $\Psi(u) = \Psi(u) \cap R_U$, where $R_U = \{r \in \Sc : \varphi(r) \in U\}$ is the set of target reports that embed into $U$.
  Finally, define $\psi(u) \in \Psi(u)$, breaking ties arbitrarily.
  If $\Psi(u)$ became empty, then leave $\psi(u)$ undefined.
\end{construction}
In \S~\ref{sec:poly-ie-consistency} we generalize this construction beyond embeddings, where we only require that $L$ indirectly elicits $\prop\ell$.
There we will see that, perhaps surprisingly, this construction recovers every possible calibrated link.

Applying Construction \ref{const:eps-thick-link} also enables one to verify the consistency of a given proposed link $\psi^*$.
For a given $\epsilon$ and norm $\|\cdot\|$, suppose one follows the routine of Construction~\ref{const:eps-thick-link} until the last step in which values for the link $\psi$ are selected.
Instead, we can simply test whether the proposed link values are contained in the valid choices, i.e., if $\psi^*(u) \in \Psi(u)$ for all $u\in\reals^d$.
If so, then the proposed link $\psi^*$ is calibrated.
\botodo{In general, how do we deal with the $\epsilon$? How do we know what $\epsilon$ to test? I guess we can gradually shrink it and see...}
\raft{A guess for $\epsilon$ was pretty clear from the top-$k$ paper, so yeah I think ``shrink and see'' makes sense.  Also probably okay to not even mention $\epsilon$ here given the last line of the paragraph.}
See \S~\ref{sec:top-k} for an illustration of this test.
On the other hand, if $\psi^*$ cannot be produced from Construction \ref{const:eps-thick-link}, then it cannot be a calibrated link.
This impossibility can be shown, for example, if there exists a point $u$ where $\Psi(u)$ is empty for all $\epsilon > 0$.

@@

Perhaps surprisingly, one can also show that \emph{every} calibrated link from a polyhedral surrogate to a discrete loss is produced by Construction \ref{const:eps-thick-link}.
This result follows from the fact, stated now, that Construction~\ref{const:eps-thick-link} with $\|\cdot\|_\infty$ is exactly enforcing $\epsilon$-separation.
From Theorem~\ref{thm:calibrated-separated}, every calibrated link $\psi$ is therefore output by the construction for some sufficiently small $\epsilon$, in the sense that $\psi$ is one of the valid choices within the link envelope.
